---
title: "p8105_hw3_to2345"
author: "Tianna Ouyang"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


```{r question1_load_dataset}
# load instacart data from p8105.datasets
library(p8105.datasets)
data("instacart")
```

* The dataset instacart has `r nrow(instacart)` rows and `r ncol(instacart)` columns;
* The variables in the datasets are: `r ls(instacart)`
* There are `r nrow(distinct(instacart, order_id))` distinct order_id in the dataset;
* There are `r nrow(distinct(instacart, user_id))` distinct user_id in the dataset;
* There are `r nrow(distinct(instacart, department))` distinct department in the dataset, for example, `r unique(pull(instacart, department))[1:5]`;
* There are `r nrow(distinct(instacart, aisle))` distinct aisel in the dataset, for example, `r unique(pull(instacart, aisle))[1:5]`.

```{r}
# How many aisles are there, and which aisles are the most items ordered from?
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n = n()
  ) %>% 
  arrange(desc(n))
```

**There are 134 aisles and most items are orderd from "fresh vegetables".**



```{r}
# Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n = n()
  ) %>% 
  filter(n > 10000) %>% 
  arrange(desc(n)) %>%
  mutate(aisle = forcats::fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_segment(aes(x = aisle, xend = aisle, y = 10000, yend = n), 
               color = "grey", 
               size = 0.7) +
  geom_point( color= "orange",
              size=2 ) +
  coord_flip() +
  theme(
          legend.position="none") +
  xlab("Aisle")+
  ylab("The number of items ordered") +
  ggtitle("The number of items ordered in each aisle")
```

**From the lollipop graph we can see clearly that "fresh vegetables" has the largest number of items ordered and "butter" has the least number number of itmes ordered among aisles with more than 10000 items ordered.**


```{r}
# Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
instacart %>% 
  filter(aisle == "baking ingredients"|aisle ==  "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(
    n = n()
  ) %>% 
  group_by(aisle) %>% 
  mutate(
    order_ranking = min_rank(desc(n))
  ) %>% 
  filter(order_ranking <= 3) %>%
  arrange(aisle, order_ranking) %>%
  knitr::kable()
```

```{r}
# Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

instacart %>% 
  filter(product_name == c("Pink Lady Apples","Coffee Ice Cream") ) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>% 
  knitr::kable(digits = 1)

```


















